%Setup%
\documentclass[12pt, letterpaper]{article}
\usepackage{math110-prep}
\title{LADR 3A}
\begin{document}
\maketitle

%Content%
\section*{The Vector Space of Linear Maps}

\begin{imp}
{3.2 Definition: linear map}
Aka linear transformation, a linear map from $V$ to $W$ is a function $T:V\to W$ 
which follows:
\begin{itemize}
  \item additivity: $(\forall u,v\in V); T(u+v)=Tu+Tv$
  \item homogeneity: $(\forall \lambda\in F, v\in V); T(\lambda v)=\lambda(Tv)$
\end{itemize}
Special: 
\begin{itemize}
  \item zero: $0v=0$. The LHS 0 is a function. The LHS 0 is the additive identity of $W$.
  \item identity: $Iv=v$. Identity map is the function that maps each element to itself.
\end{itemize}
\end{imp}
Notation: the set of all linear maps from V to W is denoted as
$\Hom(V,W)$.

\begin{imp}
{3.6 Definition: addition, scalar multiplication on $\Hom(V,W)$}
$$(S+T)(v)=Sv+Tv \text{ and } (\lambda T)(v)=\lambda(Tv)$$
\end{imp}
$\Hom(V,W)$ is a vector space!

\begin{imp}
{3.8 Definition: Product of Linear Maps}
If $T\in\Hom(U,V),S\in\Hom(V,W),u\in U$, then
$$ST\in\Hom(U,W);(ST)(u) = S(Tu)$$
\end{imp}

\begin{imp}
{3.9 Algebraic properties of products of linear maps}
\begin{itemize}
  \item associativity $(T_1T_2)T_3=T_1(T_2T_3)$
  \item identity $TI=IT=T$
  \item distributive $(S_1+S_2)T=S_1T+S_2T\text{ and }S(T_1+T_2)=ST_1+ST_2$
\end{itemize}
\end{imp}
Not commutative!

\begin{problem}
{9}
Give an example of a function $\phi:\C\to\C$ that has
additivity but not homogeneity.
\end{problem}
\begin{proof}
Let $\phi(a+bi)=a-bi$ (flip over x axis). It is easy to see that this is additive.
Let $\lambda=\cis(\pi/3)$ (60 degree CCW rotation).
\\Flip + 60 degree rotation is not equal to 60 degree rotation + flip.
\end{proof}

\begin{problem}
{10}
Let $U$ be a subspace of $V$.
$Tv = 
\begin{cases} 
  Sv & v\in U \\
  0 & v\in V, v\notin U
\end{cases}$.
Prove $T$ is not a linear map on $V$.
\end{problem}
\begin{proof}
Let $u\in U$ and $v\in V,v\notin U$. Then 
$$T(u)+T(v)=S(u)$$
$$\neq T(u+v)=S(u+v).$$
Since additivity does not hold, T is not a linear map.
\end{proof}

\begin{problem}
{14} Prove there exists $S,T\in\Hom(V,V)$ where $ST\neq TS$.
\end{problem}
\begin{proof}
We can come up with a similar solution to problem 9.
A rotation function and flip function are not commutative.
\end{proof}

\end{document}